---
title: "vis_1"
author: "My An Huynh"
date: "2024-09-26"
output: github_document
---

```{r setup, include=FALSE}
library(tidyverse)
```
 
Import weather data 
 
```{r weather}
weather_df = 
  rnoaa::meteo_pull_monitors(
    c("USW00094728", "USW00022534", "USS0023B17S"),
    var = c("PRCP", "TMIN", "TMAX"), 
    date_min = "2021-01-01",
    date_max = "2022-12-31") |>
  mutate(
    name = case_match(
      id, 
      "USW00094728" ~ "CentralPark_NY", 
      "USW00022534" ~ "Molokai_HI",
      "USS0023B17S" ~ "Waterhole_WA"),
    tmin = tmin / 10,
    tmax = tmax / 10) |>
  select(name, id, everything())
```

Making the first plot

```{r}
ggplot(weather_df, aes(x = tmin, y = tmax))+
  geom_point()
```

```{r}
weather_df |> 
ggplot(aes(x = tmin, y = tmax))+
  geom_point()
```

```{r}
ggp_weather_scatterplot = 
  weather_df |> 
  ggplot(aes(x = tmin, y = tmax))+
  geom_point()

ggp_weather_scatterplot
```

#Make fancier plots

geom_smooth gives a smooth line. (se = false) to hide standard error. 
```{r}
weather_df |> 
  ggplot(aes(x = tmin, y = tmax, color = name)) +
  geom_point(alpha = .3, size = .8) +
  geom_smooth(se = FALSE)
```
 WHere you place color = matters. See how color placed in geom_point is different. There's only one fitted smooth line instead of three. This is because placing color = name in ggplot is "global". 
```{r}
weather_df |> 
  ggplot(aes(x = tmin, y = tmax)) +
  geom_point(aes(color = name), alpha = .3, size = .8) +
  geom_smooth(se = FALSE)
```

Faceting will separate the plots. Using facet_grid(. ~ name) will place everything (.) in the rows. Name will go to column. The panels are separated by name. 

```{r}
weather_df |> 
  ggplot(aes(x = tmin, y = tmax, color = name)) +
  geom_point(alpha = .3, size = .8) +
  geom_smooth(se = FALSE) +
  facet_grid(. ~ name)

```

Let's make a somewhat more interesting scatterplot. Add another aesthetic size = so that the size of the points vary by precipitation. 
```{r}
weather_df |> 
  ggplot(aes(x = date, y = tmax, color = name, size = prcp))+
  geom_point(alpha = .3) +
  geom_smooth(se = FALSE) +
  facet_grid(. ~ name)
  
```

```{r}
weather_df |> 
  filter(name == "CentralPark_NY") |> 
  mutate(
    tmin_f = (tmin*1.8) + 32,
    tmax_f = (tmax*1.8) + 32
  ) |> 
  ggplot(aes(x = tmin_f, y = tmax_f)) +
  geom_point(alpha = .3) + 
  geom_smooth(method = "lm", se = FALSE)
  



```


